---
title: "Examples_functions"
author: "LiminChen"
date: "2024-04-15"
output: html_document
---

# Examples of functions in the stmut package
#### functions for detecting point/somatic mutation
The following are basic examples on how to run each individual function:
1. sptBClstRds()
```{r example}
library(stmut)
## basic example code
files <- list.files(path = "./inst/extdata/spotIndex", pattern = ".txt", full.names = TRUE, recursive = FALSE)
data1 <- read.csv("./inst/extdata/filtered_feature_bc.csv", header = TRUE)
data2 <- read.csv("./inst/extdata/Graph-Based.csv", header = TRUE)
path <- "./inst/extdata/tissue_positions_list.csv"

df <- sptBClstRds(files=files,data1=data1, data2=data2, path=path)

head(df[[1]]) # sptBcRds.csv
head(df[[2]]) # sptBcSummary.csv
```

2. sptMutCt()
```{r example2}
index <- read.csv("./inst/extdata/spotBC.csv", header = TRUE)
files2 <- list.files(path = "./inst/extdata/mpileup",pattern = "MpileupOutput_RNA.txt", full.names = TRUE, recursive = TRUE, include.dirs = TRUE)
df1 <- sptMutCt(index = index, files = files2)

head(df1[[1]]) # spotRdCtFinal.csv
head(df1[[2]]) # spotMutReadCount.csv
```

3. nonZeRdCts()
```{r example3}
d1 <- read.csv("./inst/extdata/spotBClster.csv", header = TRUE)
d2 <- read.csv("./inst/extdata/spotRdCtFinal.csv", header = TRUE)
d2 <- na.omit(d2)

df3 <- nonZeRdCts(df1 = d1, df2 = d2)
head(df3[[1]]) # NonZeroRdSpotIndex.csv
head(df3[[2]]) # MutantSpotIndex.csv
```

4. spotSummary()
```{r example4}
df <- read.csv("./inst/extdata/NonZeroRdSpotIndex.csv", header = TRUE)
path1 <-"./inst/extdata/mpileup/"

df4 <- spotSummary(df = df, path1 = path1)

head(df4) # NonZeroRdSpotIndex_Scored.csv
```
Column meaning of df4: <br />
column 1: Spatial barcode <br />
column 2: Spot name <br />
column 3: Cluster in Loupe.Loupe file <br />
column 4: Total reads each spot has <br />
column 5: Row position in position_lists_bc.csv <br />
column 6: Column position in position_lists_bc.csv <br />
column 7: Reference read counts <br />
column 8: Mutant read counts <br />
column 9: Sum of Ref read counts and Mut read counts (sum of col 7 and col 8) <br />
column 10: The number ( < 10) in this column indicates how many mutant reads in a spot. For example, G9 means there are 9 mutant reads in that spot. However, G10 means there are10 or more mutant reads in the spot. <br />
column 11: Proportion of # of mutant reads to Total reads (col8/col4) <br />
column 12: Proportion of # of mutant reads to sum of ref and mut reads (col8/col9) <br />
column 13: A binary column, when the spot has > 1 mutant read, it is 1, otherwise, it is 0. <br />
column 14: A score calculated based on the # of ref and mutant read counts. The higher the score is, the more likely it is a tumor spot. The table is sorted by this column. <br />
column 15: Details on which genes contain what mutations in which spots. <br />


#### functions for calling CNVs
### Examples of R functions
accStartCNR_CNS() and bulkCNVs() functions are used to generate arm-level bulk CNVs plot. <br />
wtArmMedianOne(), CtArmGenes() and cdt_filt_sort() functions are used in generating spot CNVs heatmap. Kindly refer to the case study in our manuscript FigTableScripts [here](https://github.com/limin321/stmut/blob/master/FigTableScripts/FigTables.md#spatial-spot-cnvs-figure-3) <br />

Here are examples on how to run the following functions: <br />
1. wtArmMedianOne()
```{r}
df <- read.csv("./inst/extdata/spot1_rep1.cnr", header = TRUE)
path1 <- read.csv("./inst/extdata/hg38_centromereSimple.bed", sep = "\t", header = FALSE) 

df5 <- wtArmMedianOne(data = cnr, centmere = centm)

head(df5[[1]])
head(df5[[2]])
```
2. CtArmGenes()
```{r}
cdt <- read.table("./inst/extdata/cdt.cdt", header = TRUE)
d3 <- read.table("./inst/extdata/summary.txt", sep = "\t", header = TRUE) 

df6 <- CtArmGenes(cdt = cdt, data = d3)

head(df6)
```

3. cdt_filt_sort()
```{r}
library(dplyr)
cdt <- read.table("./inst/extdata/cdt.cdt", header = TRUE)
data4 <- read.csv("./inst/extdata/CtArmGenSummary.csv",  header = TRUE)
arm <- c("1p","3p","3q","4q","5q","8q","9q","10p","10q","11q","13p","13q", "20p","20q","21q","14q","17q")
d4 <- data4 %>% filter(arms %in% arm)
genes <- d4[,"genes"]
rs <- d4[,"gene_row"]
gainLoss <- c(1,-1,1,-1,-1,1,1,-1,-1,1,-1,-1,1,1,-1,1,1)

df7 <- cdt_filt_sort(cdt = cdt,genes = genes,gainLoss = gainLoss,rs=rs)

head(df7)
```

4. cnvs()
```{r, eval=FALSE}
# data1: annotation data. A 2 column dataframe with column names as "spotName" and "annotate". Also, the annotate column must have "normal" cells or spots which will be used to center the data.
# cdt file is the weighted median cdt file generated by cnvkit.py.
# data2 is the output of wtArmMedianOne() function stored in the wtcnr/summary.txt

cnvs(data1 = data1, cdt = cdt0, data2 = data2, arm = arm, gainLoss = gainLoss) # the output is CNVs.cdt in your working directory.
```







